---
title: "Expt_b26"
author: "Max Schubert"
date: "6/19/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(stringr)
library(plyr)
require(ggplot2)
require(data.table)
require(scales)
library(gridExtra)
library(plotly)

parent_folder <- here::here('expt_b26_evolver')
source(here::here('constants.R'))
```

```{r get data}
mergebinpath <- paste(parent_folder,"merged_TC_march2019", sep = "/" )
merged_bins.dt <- data.table(fread(mergebinpath))
mergemelt.dt <- melt(merged_bins.dt, id.var=c('read_id', 'total_count', 'bc_read'), value.name='Count', variable.name='Sample')
mergemelt.dt[, SampleNum := str_replace_all(Sample, "_cutcount", "")]
mergemelt.dt = mergemelt.dt[Count > 0 ] #discard zeroes

#metadata
liblistpath <- paste(parent_folder,"b26_TC_metadata.csv", sep = "/" )
liblistdt <- data.table(fread(liblistpath))


```
#### get donors, merge to create annotated table
```{r allele_list_merge}
allelelistpath <- paste(parent_folder,"allele_annotated_list.csv", sep = "/" )
seq_annot_dt <- data.table(fread(allelelistpath))
seq_annot_dt[, bc_read := prodonor_seq]

#merge sample and seq metadata to big table
annotated_megatable <- liblistdt[mergemelt.dt, on = "SampleNum", nomatch = 0]
seq_annotated_megatable <- annotated_megatable[seq_annot_dt, on = "bc_read", nomatch = 0]

#filterstats
count_before_join = sum(annotated_megatable$Count)
count_after_join = sum(seq_annotated_megatable$Count)
filterstat = paste(as.character(count_before_join),
                   "reads in,",
                   as.character(count_after_join),
                   "reads out, or",
                   count_after_join/count_before_join*100,
                   "%", sep = " ")
print(filterstat)
```

## normalize counts to neutral allele pool, for each sample
### alternatively, calculate frequency of each seq (sec count / total count)
```{r normalize}
megatable_ratios <- seq_annotated_megatable[, mean_neutrals := mean(Count[mut_class == "Neutral"]), by='SampleNum']
megatable_ratios <- megatable_ratios[, ratio_to_neutralmean := Count/mean_neutrals] #no longer by = 'bc_read'
#same thing, but with median neutral value
megatable_ratios <- seq_annotated_megatable[, median_neutrals := as.double(median(Count[mut_class == "Neutral"])), by='SampleNum']
megatable_ratios <- megatable_ratios[, ratio_to_neutralmedian := Count/median_neutrals] #no longer by = 'bc_read'


megatable_ratios <- megatable_ratios[, total_count_sample := sum(Count), by='SampleNum']
megatable_ratios <- megatable_ratios[, seq_freq := Count/total_count_sample]

```

```{r filter_initial_no_rif}
#first two timepoints, no rif is included
early = megatable_ratios[time_point %in% c(0, 1)]
late_no_rif = megatable_ratios[expt_set == "tc"][rif == 0][time_point %in% c(2, 3, 4, 5)]
no_rif_tc <- rbind(early, late_no_rif)

#initialize to t0
no_rif_tc[,
           t0.rel.enrichment := ratio_to_neutralmedian/ ratio_to_neutralmedian[time_point==0],
           by=c("tx","nickname")] #within each replicate, for each sequence

#while we're at it, clean up the table
no_rif_tc[, c("read_id",
               "total_count",
               "bc_read",
               "options",
               "mutation",
               "barcodes",
               "drug_resistance_expected",
               "drug_sensitivity_expected"):=NULL]
```

## summarize the data, including no init version
```{r stat_summ2}
#specify current var to summarize
current_var = quote(ratio_to_neutralmedian) ####

noriftcsumm <- no_rif_tc[,
                        list(
                          Count = Count,
                          N= .N,
                          mean_var = mean(eval(current_var)),
                          sd_var= sd(eval(current_var)),
                          se_var= sd(eval(current_var)) / sqrt(.N) ),
                        by=c("nickname", "mut_class", "time_point", "ssap")]

```

# first, looking at beta
```{r plot_norif_beta}

current = noriftcsumm[ssap == "beta"] #[mut_class %in% c("Lethal", "Neutral")]

ggplot(current, aes(x=time_point, y=mean_var, color=nickname)) + 
  geom_errorbar(aes(ymin=mean_var-se_var, ymax=mean_var+se_var, alpha = .3), width=.1) +
  geom_line(aes(alpha = .3)) +
  geom_point(aes(group = nickname, alpha = .3)) +
  scale_y_continuous(trans = log10_trans() ) +#, #limit=c(0.0005, 8), 
                     #breaks = c(.0001, .001, .01, .1, 1, 10, 100, 1000, 10000)) +
  theme_bw() +
  facet_wrap(~mut_class) +
  theme(legend.position="none")

```
some lethal and deleterious seeming to drop out: gyrA and rpsL




#########OTHER CRUFT PASTED BELOW


# then, looking at PF020
```{r plot_norif_PF020}

current = noriftcsumm[ssap == "PF020"] #[mut_class %in% c("Lethal", "Neutral")]

ggplot(current, aes(x=time_point, y=mean_var, color=nickname)) + 
  geom_line(aes(alpha = .3)) +
  geom_errorbar(aes(ymin=mean_var-se_var, ymax=mean_var+se_var, alpha = .3), width=.1, color = "black") +
  geom_point(aes(group = nickname, alpha = .3, shape = as.factor(N))) +
  scale_y_continuous(trans = log10_trans(), #limit=c(0.0005, 8), 
                     breaks = c(.0001, .001, .01, .1, 1, 10, 100, 1000, 10000)) +
  theme_bw() +
  facet_wrap(~mut_class) +
  theme(legend.position="none")

```



what if we simplify to the first three timepoints?
```{r plot_no_rif_beta_short}
#could also simplify to just three classes [mut_class %in% c("Lethal", "Neutral")]
# first, looking at beta
current = noriftcsumm[ssap == "beta"][time_point %in% c(0, 1, 2)]
bet <- ggplot(current, aes(x=time_point, y=mean_var, color=nickname)) + 
  geom_errorbar(aes(ymin=mean_var-se_var, ymax=mean_var+se_var, alpha = .3), width=.1) +
  geom_line(aes(alpha = .3)) +
  geom_point(aes(group = nickname, alpha = .3)) +
  scale_y_continuous(trans = log10_trans(), #limit=c(0.0005, 8), 
                     breaks = c(.0001, .001, .01, .1, 1, 10, 100, 1000, 10000)) +
  theme_bw() +
  ggtitle("Beta, no rif") +
  facet_wrap(~mut_class) +
  theme(legend.position="none")
bet
```

#then PF020
```{r pf020}
# could [mut_class %in% c("Lethal", "Neutral")]
current = noriftcsumm[ssap == "PF020"][time_point %in% c(0, 1, 2)]
pf <- ggplot(current, aes(x=time_point, y=mean_var, color=nickname, label= N)) + 
  geom_errorbar(aes(ymin=mean_var-se_var, ymax=mean_var+se_var, alpha = .3), width=.1) +
  geom_line(aes(alpha = .3)) +
  geom_point(aes(group = nickname, alpha = .3)) +
  scale_y_continuous(trans = log10_trans(), #limit=c(0.0005, 8), 
                     breaks = c(.0001, .001, .01, .1, 1, 10, 100, 1000, 10000)) +
  theme_bw() +
  ggtitle("PF020, no rif") +
  facet_wrap(~mut_class) +
  theme(legend.position="none")
pf

#grid.arrange(bet, pf, nrow = 2)

```

### ok trying another thing. let's add back all time points, and do no normalization to neutrals
graphing the "seq freq" variable, with error bars as it's SE

```{r set_up}
#specify current var to summarize: seq freq
current_var = quote(seq_freq) ####

noriftcsumm <- no_rif_tc[,
                        list(
                          Count = Count,
                          N= .N,
                          mean_var = mean(eval(current_var)),
                          sd_var= sd(eval(current_var)),
                          se_var= sd(eval(current_var)) / sqrt(.N) ),
                        by=c("nickname", "mut_class", "time_point", "ssap")]


```

```{r plot_norif_seq_freq}

# first, looking at beta
bet_table = noriftcsumm[ssap == "beta"][time_point %in% c(0, 1, 2)]
bet <- ggplot(current, aes(x=time_point, y=mean_var, color=nickname)) + 
  geom_errorbar(aes(ymin=mean_var-se_var, ymax=mean_var+se_var, alpha = .3), width=.1) +
  geom_line(aes(alpha = .3)) +
  geom_point(aes(group = nickname, alpha = .3)) +
  scale_y_continuous(trans = log10_trans(), #limit=c(0.0005, 8), 
                     breaks = c(.0001, .001, .01, .1, 1, 10, 100, 1000, 10000)) +
  theme_bw() +
  ggtitle("Beta, no rif") +
  facet_wrap(~mut_class) +
  theme(legend.position="none")
bet
```
```{r pf020_version}
#then PF020
csp_table = noriftcsumm[ssap == "PF020"][time_point %in% c(0, 1, 2)]
pf <- ggplot(current, aes(x=time_point, y=mean_var, color=nickname)) + 
  geom_errorbar(aes(ymin=mean_var-se_var, ymax=mean_var+se_var, alpha = .3), width=.1) +
  geom_line(aes(alpha = .3)) +
  geom_point(aes(group = nickname, alpha = .3)) +
  scale_y_continuous(trans = log10_trans(), #limit=c(0.0005, 8), 
                     breaks = c(.0001, .001, .01, .1, 1, 10, 100, 1000, 10000)) +
  theme_bw() +
  ggtitle("PF020, no rif") +
  facet_wrap(~mut_class) +
  theme(legend.position="none")
pf
#grid.arrange(bet, pf, nrow = 2)

```

### i think it's time to apply the linear model, to seq freq variable
setup:
```{r setup}
#function for linear modeling
lm_group <- function(dt_subset, byval) {
  lmod <- lm(log10(mean_var) ~ time_point, data= dt_subset[is.finite(mean_var)])
  return(list(intercept= lmod$coefficients[1], slope= lmod$coefficients[2], mut_class= dt_subset$mut_class[1]))
}

#run linear model on beta and csprecT data
full_table = noriftcsumm[(ssap == "beta")|(ssap == "PF020")][time_point %in% c(0, 1, 2)]
full_table_slopes <- full_table[,
  lm_group(.SD,.BY),
  by=c('nickname', "ssap")]

```


#ok then actually plot the slopes
first with beta:
```{r}
fig_bet <- ggplot(full_table_slopes[ssap == "beta"],
                  aes(x=mut_class, y=slope,
                      color = mut_class, group = nickname)) +
  geom_hline(yintercept=0, linetype=2, size=0.5) + #baseline no slope
  geom_jitter(width = 0.2, alpha = 0.7, size = 2) +
  #annotate points below =0.05
  geom_text(aes(label = nickname), vjust = -1.1,
            data = subset(full_table_slopes[ssap == "beta"],
                                        slope < -.05)) +
  geom_boxplot(aes(group = mut_class), alpha = 0.7,
               outlier.shape = NA) +
  #facet_grid(~mut_class, scales='free_x', switch = 'both') +
  theme_linedraw() +
  labs(y = "RLR slopes (Beta, Rif0)") +
  #scale_x_discrete(labels = x_labels) +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1,
                               size = fig_font_size),
    axis.text.y = element_text(size = fig_font_size),
    axis.title.y = element_text(size = fig_title_font_size),
    axis.title.x = element_blank(),
    strip.background=element_blank(),
    panel.grid.minor=element_blank(),
    panel.grid.major.x=element_blank(),
    panel.border=element_blank(),
    axis.line=element_line(),
    legend.position = 'none')

fig_bet

```


then with csprecT
```{r}
fig_csp <- ggplot(full_table_slopes[ssap == "PF020"],
                  aes(x=mut_class, y=slope,
                      color = mut_class, group = nickname)) +
  geom_hline(yintercept=0, linetype=2, size=0.5) + #baseline no slope
  geom_jitter(width = 0.2, alpha = 0.7, size = 2)+
  geom_boxplot(aes(group = mut_class), alpha = 0.7,
               outlier.shape = NA) +
  #annotate points below =0.05
  geom_text(aes(label = nickname), vjust = -1.1,
            data = subset(full_table_slopes[ssap == "PF020"],
                                        slope < -.5)) +
  #facet_grid(~mut_class, scales='free_x', switch = 'both') +
  theme_linedraw() +
  labs(y = "RLR slopes (CsprecT, Rif0)") +
  #scale_x_discrete(labels = x_labels) +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1,
                               size = fig_font_size),
    axis.text.y = element_text(size = fig_font_size),
    axis.title.y = element_text(size = fig_title_font_size),
    axis.title.x = element_blank(),
    strip.background=element_blank(),
    panel.grid.minor=element_blank(),
    panel.grid.major.x=element_blank(),
    panel.border=element_blank(),
    axis.line=element_line(),
    legend.position = 'none')

fig_csp
```
```{r}
grid.arrange(fig_bet, fig_csp, nrow = 2)
```








ratio to neutral median, rather than mean
```{r plot_norif_seq_freq_median}
# first, looking at beta
current = noriftcsumm[ssap == "beta"][time_point %in% c(0, 1, 2, 3)][mut_class %in% c("Lethal", "Neutral", "Drug_resistance")]
bet <- ggplot(current, aes(x=time_point, y=mean_rationeutmed, color=nickname)) + 
  geom_errorbar(aes(ymin=mean_rationeutmed-se_rationeutmed, ymax=mean_rationeutmed+se_rationeutmed, alpha = .3), width=.1) +
  geom_line(aes(alpha = .3)) +
  geom_point(aes(group = nickname, alpha = .3)) +
  scale_y_continuous(trans = log10_trans(), #limit=c(0.0005, 8), 
                     breaks = c(.0001, .001, .01, .1, 1, 10, 100, 1000, 10000)) +
  theme_bw() +
  ggtitle("Beta, no rif") +
  facet_wrap(~mut_class) +
  theme(legend.position="none")
#then PF020
current = noriftcsumm[ssap == "PF020"][time_point %in% c(0, 1, 2, 3)][mut_class %in% c("Lethal", "Neutral", "Drug_resistance")]
pf <- ggplot(current, aes(x=time_point, y=mean_rationeutmed, color=nickname)) + 
  geom_errorbar(aes(ymin=mean_rationeutmed-se_rationeutmed, ymax=mean_rationeutmed+se_rationeutmed, alpha = .3), width=.1) +
  geom_line(aes(alpha = .3)) +
  geom_point(aes(group = nickname, alpha = .3)) +
  scale_y_continuous(trans = log10_trans(), #limit=c(0.0005, 8), 
                     breaks = c(.0001, .001, .01, .1, 1, 10, 100, 1000, 10000)) +
  theme_bw() +
  ggtitle("PF020, no rif") +
  facet_wrap(~mut_class) +
  theme(legend.position="none")

grid.arrange(bet, pf, nrow = 2)

```











# the end

##### session info, for reproducibility
```{r sesh, echo = FALSE, include = FALSE}
sessionInfo()
```

