---
title: "Fig_3CD"
author: "Max Schubert"
date: "6/18/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(stringr)
library(plyr)
require(ggplot2)
require(data.table)
require(scales)
library(grid)
library(gridExtra)
library(cowplot)
library(here)
library(ggpubr)
library(here)

parent_folder <- here::here('fig_3CD')
source(here::here('constants.R'))
```

```{r mergebins_liblist}
mergebinpath <- paste(parent_folder,"merged_TC_march2019", sep = "/" )
merged_bins.dt <- data.table(fread(mergebinpath))
mergemelt.dt <- melt(merged_bins.dt, id.var=c('read_id', 'total_count', 'bc_read'), value.name='Count', variable.name='Sample')
mergemelt.dt[, SampleNum := str_replace_all(Sample, "_cutcount", "")]
rm(merged_bins.dt) #remove original

#metadata
liblistpath <- paste(parent_folder,"b26_TC_metadata.csv", sep = "/" )
liblistdt <- data.table(fread(liblistpath))

#### get donors, merge to create annotated table
allelelistpath <- paste(parent_folder,"allele_annotated_list.csv", sep = "/" )
seq_annot_dt <- data.table(fread(allelelistpath))
seq_annot_dt[, bc_read := prodonor_seq]

#merge sample and seq metadata to big table
annotated_megatable <- liblistdt[mergemelt.dt, on = "SampleNum", nomatch = 0]
rm(mergemelt.dt) #remove original, large file
seq_annotated_megatable <- annotated_megatable[seq_annot_dt, on = "bc_read", nomatch = 0]

#print statement about merge efficiency
num_merged = sum(annotated_megatable$Count)
rm(annotated_megatable) #remove original, millions of rows
num_annot = sum(seq_annotated_megatable$Count)
pct_annot = num_annot/num_merged
num_merged = as.character(num_merged)
num_annot = as.character(num_annot)
pct_annot = as.character(pct_annot)




print(paste(num_annot, "reads annotated of",
            num_merged, "total reads, or",
            100*as.numeric(pct_annot), "%", sep = " "))
```

```{r PULL_NEW_DATA}
mergebinpath2 <- paste(parent_folder,"H6_merged_bins", sep = "/" )
merged_bins2.dt <- data.table(fread(mergebinpath2))
mergemelt2.dt <- melt(merged_bins2.dt, id.var=c('read_id', 'total_count', 'bc_read'), value.name='Count', variable.name='Sample')
mergemelt2.dt[, SampleNum := str_replace_all(Sample, "_cutcount", "")]
rm(merged_bins2.dt) #remove original

#metadata
liblistpath2 <- paste(parent_folder,"liblistH6.csv", sep = "/" )
liblistdt2 <- data.table(fread(liblistpath2))
liblistdt2[,SampleNum := as.character(SampleNum)]
#liblistdt2[,samplenum:= NULL]


allelelistpath <- paste(parent_folder,"allele_annotated_list.csv", sep = "/" )
seq_annot_dt <- data.table(fread(allelelistpath))
seq_annot_dt[, bc_read := prodonor_seq]

#merge sample and seq metadata to big table
annotated_megatable2 <- liblistdt2[mergemelt2.dt, on = "SampleNum", nomatch = 0]
rm(mergemelt2.dt) #remove original, millions of rows
seq_annotated_megatable2 <- annotated_megatable2[seq_annot_dt, on = "bc_read", nomatch = 0]

#print statement about merge efficiency
num_merged = sum(annotated_megatable2$Count)
rm(annotated_megatable2) #remove original, millions of rows
num_annot = sum(seq_annotated_megatable2$Count)
pct_annot = num_annot/num_merged
num_merged = as.character(num_merged)
num_annot = as.character(num_annot)
pct_annot = as.character(pct_annot)

print(paste(num_annot, "reads annotated of",
            num_merged, "total reads, or",
            100*as.numeric(pct_annot), "%", sep = " "))
```

```{r add new data for timepoint 1 to old}
#create "before" table for old data, new, with beta only
before = seq_annotated_megatable[time_point == 1][tx %in% c(1,2,3)]
before2 = seq_annotated_megatable2[time_point == 1][tx %in% c(1,2,3)]


both_before = rbind(before, before2, fill = TRUE)
#TDO TODO need to trim table down to smaller set of things, if I'm going to combine my timepoint, etc
both_before[,Count_both := sum(Count), by = c('prodonor_seq', 'time_point', 'tx')]

before = both_before
```

```{r segment}
#make versions of "before" data assigned to each value of rif, so i can segment on rif
before12.5 = before
before12.5$rif = 12.5
before25 = before
before25$rif = 25
before50 = before
before50$rif = 50
before100 = before
before100$rif = 100

after = seq_annotated_megatable[(expt_set == "binary")]
after[,Count_both := Count]
before_after_1  <- rbind(before12.5, before25, before50, before100, after, fill = TRUE)
before_after_1 <- before_after_1[expt_set != "tc_2"]


#rm(before, after, before12.5, before25, before50, before100) #remove transitional tables

#while we're at it, clean up the table
before_after_1[, c("Library Id","read_id","Library Name","ssap","operation",
               "total_count",
               "bc_read",
               "options",
               "mutation",
               "barcodes",
               "drug_resistance_expected",
               "drug_sensitivity_expected"):=NULL]
```

## calculate frequency of each seq (sec count / total count)
```{r calculate frequencies}
before_after_1[, total_count_sample := sum(Count_both), by=c('time_point','tx', 'rif')]
#important: pseudocount of one given to undetected sequences AFTER calculating totals
before_after_1[Count_both == 0, Count_both := 1]

before_after_1[, seq_freq := Count_both/total_count_sample]
```
at this point, the analysis branches. here i'm examining the ratios of different alleles in rif platings of different conc, to their "initial" frequency in the t = 1 samples
 
#Rif
```{r ratio_summarize}
before_after_1[, enrichment.post.rif := seq_freq / seq_freq[time_point==1],
      by=c("tx","nickname","rif")] #within each replicate, for each sequence

into_summ = before_after_1[time_point == 3] #[Count > 5] #throwing out count filter!
binary_summ <- into_summ[,
                        list(
                          Count = mean(Count),
                          N= .N,
                          number_undetected = sum(Count_both == 1),
                          mean_seq_freq = mean(seq_freq),
                          median_seq_freq = median(seq_freq),
                          mean_enrichment = mean(enrichment.post.rif),
                          median_enrichment = median(enrichment.post.rif),
                          max_enrichment = max(enrichment.post.rif),
                          logmean_enrichment = 10^mean(log10(enrichment.post.rif)),
                          logsd_enrichment = sd(log10(enrichment.post.rif)),
                          sd_freq = sd(seq_freq),
                          se_freq = sd(seq_freq) / sqrt(.N),
                          sd_enrichment= sd(enrichment.post.rif),
                          se_enrichment= sd(enrichment.post.rif) / sqrt(.N),
                          log_se_enrichment= 10^sd(log10(enrichment.post.rif)) / sqrt(.N)
                          ), by=c("nickname", "mut_class", "rif",
                                  "color_cat", "rpoB allele type",
                                  "aa_position", "aa_ref", "aa_alt")]

```

#prepare supplemental table 1A, list of figure 2C results for all alleles
```{r rif_allele_summary_supp}
output_table_1A <- binary_summ[rif == 12.5]
output_table_1A[, c("rpoB allele type", "mut_class", #color_cat
                 "Count", "N", "mean_seq_freq", "median_seq_freq",
                 "median_enrichment", "max_enrichment","logmean_enrichment", 
                 "sd_freq", "se_freq", "sd_enrichment","color_cat",
                 "log_se_enrichment", "aa_position", "aa_ref", "aa_alt", "rif"):=NULL]

#rank highest to lowest by mean enrichment
setorder(output_table_1A, 'mean_enrichment')

#output table
output_to = paste(here::here(), 'supplemental_materials', 'supplemental_table_1A.csv', sep = "/")
fwrite(output_table_1A,output_to)

```

# Figure 3C
```{r figure3c_prep, fig.width=6, fig.height=3, units='in', dpi=300}

color_vals = c('#00BA38', #green
               '#619CCF', #dark blue
               #'#F564E3', #purple
               '#F8766D') #, #red
               #'#B79F00'# #gold

#addline_format <- function(x,...){gsub(' ','\n',x)}

current = binary_summ[rif == 12.5] #rif concentration for binary plot here


# lower limit, unfilled/filled, combine del/lethal, neutral/unrelated
#also plot all rif resistance and rpoB alleles together
current_combined = copy(current)
current_combined[mut_class == 'Deleterious', mut_class := 'Lethal']
current_combined[mut_class == 'Lethal', mut_class := 'Lethal,\nDeleterious']
current_combined[mut_class == 'Lethal,\nDeleterious', color_cat := 1]

current_combined[mut_class == 'Unrelated Resistance Alleles', mut_class := 'Neutral']
current_combined[mut_class == 'Neutral', mut_class := 'Neutral,\nUnrelated Resistance']
current_combined[mut_class == 'Neutral,\nUnrelated Resistance', color_cat := 2]

#plot all rpoB alleles together
current_combined[mut_class == 'Rifampicin Resistance', mut_class := 'rpoB\nMissense']
current_combined[mut_class == 'rpoB Allele, unknown effect', mut_class := 'rpoB\nMissense']
current_combined[mut_class == 'rpoB\nMissense', color_cat := 3]

#reorder
current_combined[, mut_class := factor(
  mut_class,
  levels=unique(mut_class)[c(1,2,3)],
  labels=unique(mut_class)[c(1,2,3)])]
  
#current_combine_lethal[, mut_class := factor(mut_class, levels=levels(mut_class)[c(3:6)])]
```
#execute figure 3C
```{r fig3C, fig.width=6, fig.height=3, units='in', dpi=300}
fig3c <- ggplot(current_combined, aes(
    x=reorder(nickname,logmean_enrichment), 
    y=pmax(1e-4, logmean_enrichment), 
    color= as.factor(color_cat),
    fill= as.factor(color_cat))) + 
  geom_linerange(aes(
    ymin=pmax(1e-4,logmean_enrichment-se_enrichment), 
    ymax=pmax(1e-4,logmean_enrichment+se_enrichment)),
    color="black", 
    alpha = 0.3) +  
  geom_point(data=current_combined[number_undetected == 0], size = 2, shape=21) +
  geom_point(data=current_combined, size = 2, shape=21, fill=NA) +
  geom_text(data=current_combined[number_undetected > 0 & number_undetected < 3], label="\u25D3", size=4.7, family = "Arial Unicode MS") +
  geom_point(data=current_combined[number_undetected == 3], size = 2, shape=21, fill='white') +
  scale_colour_manual(values = color_vals[1:5]) +
  scale_fill_manual(values = color_vals[1:5]) +
  scale_y_continuous(
    trans = log_trans(base=10),
    breaks = c(.0001, .001, .01, .1, 1, 10, 100),
    labels = c('\u2265 -4', log10(c(.001,.01,.1, 1, 10, 100))),
    limits=10^(c(-4.2,1.6)), 
    expand=c(-0.01,0)) +
  scale_x_discrete(expand=c(0.16,0.16)) +
  facet_grid(~mut_class, scales='free_x', switch = 'both') + #, scales='free_x', space='free_x', switch='both') +
  labs(
    x = "Allele",
    y = "Allelic enrichment\n(log10)",
    color = "Allele type") + 
  theme_linedraw() +
  geom_hline(yintercept=1, linetype=2) +
  theme(text = element_text(size=fig_font_size)) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = fig_font_size),
    axis.title.y = element_text(size = fig_title_font_size),
    axis.title.x = element_blank(),
    axis.ticks.x=element_blank(),
    strip.background=element_blank(), strip.placement='outside',
    strip.text=element_text(color='black'), 
    panel.spacing=unit(0,'pt'),
    panel.border=element_blank(),
    panel.grid.major.x=element_blank(),
    panel.grid.minor.x=element_blank(),
    axis.line=element_line(),
    strip.text.x = element_text(size = fig_font_size),
    
    legend.position = 'none')

fig3c
```
#save figure 3C
```{r save_3C}
#ggsave('fig2c.pdf', plot=fig2c, width=6, height=3, units='in', dpi=300)

ggsave(filename='fig_3C_new2.pdf', plot=fig3c, width=5, height=2.5, units='in', dpi=300, device=cairo_pdf)

```

#prepare supplemental table 1B, list of figure 3CD results for rif alleles
```{r rif_allele_summary_supp}
output_table <- binary_summ[mut_class %in% c("Rifampicin Resistance",
                                             "rpoB Allele, unknown effect")]
output_table[, c("rpoB allele type", "mut_class", #color_cat
                 "Count", "N", "mean_seq_freq", "median_seq_freq",
                 "median_enrichment", "max_enrichment",
                 "min_enrichment", "logmean_enrichment", 
                 "sd_freq", "se_freq", "sd_enrichment",
                 "log_se_enrichment", "aa_position", "aa_ref", "aa_alt"):=NULL]

#rank highest to lowest by mean enrichment
output_table[,rank_per_conc := rank(-mean_enrichment), by = "rif"]
output_to = paste(parent_folder, 'rif_binary_out.csv', sep = "/")
fwrite(output_table,output_to)

#output all alleles
unique(binary_summ$nickname)
```

#setup figure 3D
```{r plot_concentration_change, fig.width=3, fig.height=3, units='in', dpi=300}


fig3d <- ggplot(binary_summ[mut_class=="Rifampicin Resistance" | nickname == 'rpoB_C_1576_G'], 
    aes(x=rif, y=median_enrichment, group=nickname )) + 
  geom_point( size = 2, alpha  = 0.7, stroke = 0) +
  geom_line( alpha = 0.7) +
  scale_y_continuous(trans = log_trans(base=10),
                     breaks = c(.000001, .00001, .0001, .001, .01, .1, 1, 10, 100),
                     labels = log10(c(.000001, .00001, .0001, .001, .01, .1, 1, 10, 100)),
                     limits=10^(c(-5.2,2))) +
  scale_x_continuous(trans = log_trans(base=2),
                     breaks = c(12.5, 25, 50, 100)) +
  labs(
    x = "Rifampicin Concentration\n(µg/mL)",
    y = "log10 Enrichment",
    color = "") + 
  geom_hline(yintercept=1, linetype=2) +
  theme_linedraw() +
  theme(
    axis.text.x = element_text(size = fig_font_size),
    axis.text.y = element_text(size = fig_font_size),
    axis.title.y = element_text(size = fig_title_font_size),
    axis.title.x = element_text(size = fig_title_font_size),
    panel.grid.minor.x=element_blank(),
    panel.grid.major.x=element_blank(),
    panel.grid.minor.y=element_blank(),
    panel.border=element_blank(),
    axis.line=element_line(),
    legend.position = 'none') 

fig3d
```


```{r save_fig3D}
ggsave(here::here('fig_3D.pdf'), plot=fig3d, width=3, height=3, units='in', dpi=300, device = cairo_pdf)

```








######## cruft to be deleted...
# older version of 2c plot
```{r grouped_summarized, fig.width=6, fig.height=3, units='in', dpi=300}

color_vals = c('#00BA38', #green
               '#619CCF', #dark blue
               '#F564E3', #purple
               '#F8766D', #red
               '#B79F00') #gold

addline_format <- function(x,...){gsub(' ','\n',x)}

current = binary_summ[rif == 12.5] #rif concentration for binary plot here

current[, mut_class := factor(
  mut_class,
  levels=unique(mut_class)[c(1,2,3,4,6,5)],
  labels=addline_format(unique(mut_class)[c(1,2,3,4,6,5)]))]
  
# lower limit, unfilled/filled
current_combine_lethal = copy(current)
current_combine_lethal[mut_class == 'Deleterious', mut_class := 'Lethal']
current_combine_lethal[mut_class == 'Lethal', mut_class := 'Lethal/Deleterious']
current_combine_lethal[mut_class == 'Lethal/Deleterious', color_cat := 2]
current_combine_lethal[, mut_class := factor(mut_class, levels=levels(mut_class)[c(7,3:6)])]

fig2c <- ggplot(current_combine_lethal, aes(
    x=reorder(nickname,logmean_enrichment), 
    y=pmax(1e-4, logmean_enrichment), 
    color= as.factor(color_cat),
    fill= as.factor(color_cat))) + 
  geom_linerange(aes(
    ymin=pmax(1e-4,logmean_enrichment-iqr_enrichment), 
    ymax=pmax(1e-4,logmean_enrichment+iqr_enrichment)),
    color="black", 
    alpha = 0.3) +  
  geom_point(data=current_combine_lethal[number_undetected == 0], size = 2, shape=21) +
  geom_point(data=current_combine_lethal, size = 2, shape=21, fill=NA) +
  geom_text(data=current_combine_lethal[number_undetected > 0 & number_undetected < 3], label="\u25D3", size=4.7, family = "Arial Unicode MS") +
  geom_point(data=current_combine_lethal[number_undetected == 3], size = 2, shape=21, fill='white') +
  scale_colour_manual(values = color_vals[1:5]) +
  scale_fill_manual(values = color_vals[1:5]) +
  scale_y_continuous(
    trans = log_trans(base=10),
    breaks = c(.001, .01, .1, 1, 10, 100),
    labels = c('\u2265 -3', log10(c(.01, .1, 1, 10, 100))),
    limits=10^(c(-4.2,1.6)),
    expand=c(-0.01,0)) +
  scale_x_discrete(expand=c(0.16,0.16)) +
  facet_grid(~mut_class, scales='free_x', space='free_x', switch='both') +
  labs(
    x = "Allele",
    y = "Allelic enrichment (log10)\n 12.5 µg/mL Rifampicin",
    color = "Allele type") + 
  theme_linedraw() +
  geom_hline(yintercept=1, linetype=2) +
  theme(text = element_text(size=fig_font_size)) +
  theme(
    axis.text.x= element_blank(),
    axis.ticks.x=element_blank(),
    strip.background=element_blank(), strip.placement='outside',
    strip.text=element_text(color='black'), 
    panel.spacing=unit(0,'pt'),
    panel.border=element_blank(),
    panel.grid.major.x=element_blank(),
    panel.grid.minor.x=element_blank(),
    axis.line=element_line(),
    legend.position = 'none')

fig2c

#ggsave('fig2c.pdf', plot=fig2c, width=6, height=3, units='in', dpi=300)

ggsave(filename='fig2c.pdf', plot=fig2c, width=6, height=3, units='in', dpi=300, device=cairo_pdf)

```


#prepare supplemental table 1B, list of figure 3CD results for rif alleles
```{r rif_allele_summary_supp}
output_table <- binary_summ[mut_class %in% c("Rifampicin Resistance",
                                             "rpoB Allele, unknown effect")]
output_table[, c("rpoB allele type", "mut_class", #color_cat
                 "Count", "N", "mean_seq_freq", "median_seq_freq",
                 "median_enrichment", "max_enrichment",
                 "min_enrichment", "logmean_enrichment", 
                 "sd_freq", "se_freq", "sd_enrichment",
                 "log_se_enrichment", "aa_position", "aa_ref", "aa_alt"):=NULL]

#rank highest to lowest by mean enrichment
output_table[,rank_per_conc := rank(-mean_enrichment), by = "rif"]
output_to = paste(parent_folder, 'rif_binary_out.csv', sep = "/")
fwrite(output_table,output_to)

#output all alleles
unique(binary_summ$nickname)
```


######### deprecated: plotting both as one panel

```{r fig2bc, fig.width=6, fig.height=2, units='in', dpi=300}

total_width=figure_width
relative_bc_width = 2
total_height=total_width/(relative_bc_width+1)

fig2b_comb <- fig2b + theme(plot.margin = unit(c(0, 0, 0, 0), "pt"))
fig2c_comb <- fig2c + labs(y='') + theme(plot.margin = unit(c(0, 0, 0, 0), "pt"))
grid.newpage()
fig2bc <- plot_grid(
  plotlist=list(fig2b_comb, fig2c_comb),
  align = 'hv',
  axis='tblr',
  rel_widths = c(relative_bc_width,1))

fig2bc

ggsave(here::here('fig2bc.pdf'), plot=fig2bc, width=total_width, height=total_height, units='in', dpi=300)


```
