---
title: "fig_4X"
author: "Max Schubert"
date: "10/10/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
#library(stringr)
library(plyr)
library(gridExtra)
require(ggplot2)
require(data.table)
require(scales)
library(ggpubr)
library(plotly)
#dan's suggestion for plotting:
library(ggrepel)

#for reading sam files
library(Rsamtools) # first: BiocManager::install("Rsamtools")


parent_folder = "/Users/Max_Schubert/dropbox_hms/active_seq_reads/NovNextseq_again/FC_04398/Unaligned_1234_PF_mm1/Data/sam_without_pear/" #temporary, will relocate to git here::here('fig_4x'
parent_folder = here::here('fig_4B_lengths/')
source(here::here('constants.R'))
```
### here's the code for the "tlen" column of a sam/bam file
##### there are two rows for each read pair, need to be filtered down to one by removing tlen less than zero

```{r}
#sample = "S1"

#sample_path = paste(parent_folder, sample, const, sep = "")
#lens = data.table(fread(sample_path))
#pos_lens <- lens[ V1 > 0]
#filter_lens <- pos_lens[V1 < 400]

const = "_tlen"
mk_lens_table <- function(x,...){
  sample_path = paste(parent_folder, x, const, sep = "")
  lens = data.table(fread(sample_path))
  pos_lens <- lens[ V1 > 0]
  filter_lens <- pos_lens[V1 < 400]
  return(filter_lens)
}

s1 = mk_lens_table("S1")
s2 = mk_lens_table("S2")
s3 = mk_lens_table("S3")

s1[, sample := 'WT']
s2[, sample := 'Strain 1']
s3[, sample := 'Strain 2']

combined <- rbind(s1, s2, s3)

rm(s1,s2,s3) #remove individual tables
#hist(combined$V1, breaks = 100) #simple histogram
```

```{r plot_lengths}

fig_4b_lengths <- ggplot(combined, aes(V1, color = sample)) +
  geom_freqpoly(binwidth = 10 ) + 
  theme_linedraw() +
  labs(
    x = "Donor Length",
    y = "Frequency",
    color = "Sample") +
  theme(
    panel.grid.minor.y =element_blank(),
    panel.grid.minor.x=element_blank(),
    axis.line = element_line(),
    panel.border=element_blank() )

fig_4b_lengths

ggsave('fig_4B_lengths.pdf', plot=fig_4b_lengths, width=4, height=3, units='in', dpi=300, useDingbats = FALSE)

```






###below was alternate code where i was using full bam files
```{r get_bam}

# as per https://gist.github.com/davetang/6460320

sample = "S1_aln.bam"
sample_path = paste(parent_folder, sample, sep = "")
sample_bam <- scanBam(sample_path)


#names of the BAM fields
names(sample_bam[[1]])

#distribution of BAM flags
table(sample_bam[[1]]$flag)

#function for collapsing the list of lists into a single list
#as per the Rsamtools vignette
.unlist <- function (x){
   ## do.call(c, ...) coerces factor to integer, which is undesired
   x1 <- x[[1L]]
   if (is.factor(x1)){
      structure(unlist(x), class = "factor", levels = levels(x1))
   } else {
      do.call(c, x)
   }
}

#store names of BAM fields
bam_field <- names(sample_bam[[1]])

#go through each BAM field and unlist
list <- lapply(bam_field, function(y) .unlist(lapply(sample_bam, "[[", y)))


#try making into data.table in place
#setDT(list, giveNames=TRUE, keep.rownames=FALSE)

#store as data frame
bam_df <- do.call("DataFrame", list)
names(bam_df) <- bam_field

setDT(bam_df) #doesn't work

#report dimensions
dim(bam_df)

#calculate length
bam_df$aln_len <- abs(bam_df$mpos - bam_df$pos)

#mk hist
hist(bam_df$aln_len, breaks = 100)
#bamFile

#rm NA values
bam_df <- bam_df[complete.cases(bam_df[ , 14]),]

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
