---
title: "fig_S1A_growth"
author: "Max Schubert"
date: "8/7/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup}
#load libraries
library(data.table)
library(stringr)
library(plyr)
library(gridExtra)
require(ggplot2)
require(data.table)
require(scales)
library(ggpubr)
library(RColorBrewer)


parent_folder = here::here('supplemental_figures/fig_S1A')
```


```{r get_data}
#paths to data tables, output by matlab
base_path = parent_folder

pr1a = paste(base_path,"PC1A.csv", sep = "/" )
pr1a_2 = paste(base_path,"PC1A_2.csv", sep = "/" )
pr1b = paste(base_path,"PC1B.csv", sep = "/" )
pr1b_2 = paste(base_path,"PC1B_2.csv", sep = "/" )
pr2a = paste(base_path,"PC2A.csv", sep = "/" )
pr2a_2 = paste(base_path,"PC2A_2.csv", sep = "/" )
pr2b = paste(base_path,"PC2B.csv", sep = "/" )
pr2b_2 = paste(base_path,"PC2B_2.csv", sep = "/" )

#todo: get plate map, tx annotations
#################
platemap = paste(base_path, "platemaps.csv", sep = "/")
txs = paste(base_path, "txs.csv", sep = "/")

#make into datatables
pr1a.dt = data.table(fread(pr1a ))
pr1a_2.dt = data.table(fread(pr1a_2 ))
pr1b.dt = data.table(fread(pr1b ))
pr1b_2.dt = data.table(fread(pr1b_2 ))
pr2a.dt = data.table(fread(pr2a ))
pr2a_2.dt = data.table(fread(pr2a_2 ))
pr2b.dt = data.table(fread(pr2b ))
pr2b_2.dt = data.table(fread(pr2b_2 ))

################ todo
platemap.dt = data.table(fread(platemap))
txmap.dt = data.table(fread(txs))

```


```{r reformat_data}

#label all the plates, including replicates, rbind all data together
pr1a.dt$set = 1
pr1a_2.dt$set = 1
pr1b.dt$set = 1
pr1b_2.dt$set = 1
pr2a.dt$set = 2
pr2a_2.dt$set = 2
pr2b.dt$set = 2
pr2b_2.dt$set = 2
pr1a.dt$bio_rep = 'a'
pr1a_2.dt$bio_rep = 'a'
pr1b.dt$bio_rep = 'b'
pr1b_2.dt$bio_rep = 'b'
pr2a.dt$bio_rep = 'a'
pr2a_2.dt$bio_rep = 'a'
pr2b.dt$bio_rep = 'b'
pr2b_2.dt$bio_rep = 'b'
pr1a.dt$tech_rep = 1
pr1a_2.dt$tech_rep = 2
pr1b.dt$tech_rep = 1
pr1b_2.dt$tech_rep = 2
pr2a.dt$tech_rep = 1
pr2a_2.dt$tech_rep = 2
pr2b.dt$tech_rep = 1
pr2b_2.dt$tech_rep = 2

combined = rbind(pr1a.dt,
                 pr1a_2.dt,
                 pr1b.dt,
                 pr1b_2.dt,
                 pr2a.dt,
                 pr2a_2.dt,
                 pr2b.dt,
                 pr2b_2.dt)

#merge platemaps onto combined
#you expect tx = 0 to drop, these are blank wells anyway
merged_1 <- combined[platemap.dt, on = c("well","bio_rep"), nomatch = 0]
#metadata[,c( "oligo","seq","soft_desc","recov") :=NULL]

#merge metadata onto data on tx on plate, well
total_merged <- merged_1[txmap.dt, on = c("tx", "set"), nomatch = 0]

```

```{r normalize}
### normalize to gfp control
#important: the "by" statement here is critical
total_merged[, mean_GFP := mean(mu_hourly[desc == "gfp/yfp"]), by=c('set', 'bio_rep','tech_rep')]
total_merged[, mu_normalGFP := mu_hourly/mean_GFP, by=c('set', 'bio_rep','tech_rep')]
#also normalize to scrambled? why not? might drop GFP if it's weird
total_merged[, mean_scrambled := mean(mu_hourly[desc == "scramble/empty"]), by=c('set', 'bio_rep','tech_rep')]
total_merged[, mu_normalscrambled := mu_hourly/mean_scrambled, by=c('set', 'bio_rep','tech_rep')]

#alternative version: take means across technical reps
#mean gfp per plate
bio_reps_only = total_merged[, mean_GFP := mean(mu_hourly[desc == "gfp/yfp"]), by=c('set', 'bio_rep', 'tech_rep')]
bio_reps_only[, mu_normalGFP := mu_hourly/mean_GFP] #normalize each measurement to GFPavg in the plate

bio_reps_only[, mean_tech_reps := mean(mu_normalGFP), by = c('set', 'bio_rep','well')]

bio_reps_only = bio_reps_only[tech_rep == 1]

#check whether technical reps correlate?
#is technical noise more significant than well/well variation? or vice versa?

```

```{r plot}
### version paired

#fix the x labels
addline_format <- function(x,...){
  gsub(',','\n',x)
}
x_labels <- addline_format(c("rpoB 1721,C>A",
                             "rpoB 1546,G>T",
                             "rpoB 1714,A>C",
                             "gyrA 248,C>T",
                             "rpoB 1594,delete12bp",
                             "Scrambled sequence",
                             "rpoB 443,A>T",
                             "gyrA 248,C>T (Reverse-complement)",
                             "rpoB 1547,A>G",
                             "lacZ,KO",
                             "rpoB 1538,A>C",
                             "GFP Control",
                             "rpoB 1535,C>T"))

into_plot = total_merged
ggplot(into_plot, aes(x=reorder(desc,mu_normalGFP),y=mu_normalGFP, color = as.factor(set))) +
  geom_boxplot(position = position_dodge(0.7), width = 0.5) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1, size = 12), # 
        axis.text.y = element_text(size = 12)) +
  scale_y_continuous(breaks = c(.5, 1, 1.5, 2)) +
  labs(
    x = "Retron Donor Sequence",
    y = "Growth Rate (normalized to GFP control)",
    title = "Effect of retron vector on growth rate",
    color = "Retron vector") +
  scale_x_discrete(labels = x_labels)


```
```{r summarized_plot}
into_plot = total_merged
ggplot(into_plot, aes(x=interaction(Vector_type),y=mu_normalGFP, color = Vector_type)) +
  #geom_jitter(width = .2, size = 3) +
  geom_violin(trim=FALSE) +
  geom_boxplot(position = position_dodge(0.7), width = 0.1, alpha = 0.7,
               outlier.shape = NA) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1, size = 12), # 
        axis.text.y = element_text(size = 12)) +
  scale_y_continuous(breaks = c(.7,.8,.9, 1, 1.1)) +
  labs(
    x = "Retron Donor Sequence",
    y = "Growth Rate (normalized to GFP control)",
    title = "Effect of retron vector on growth rate",
    color = "Retron vector") 
  #scale_x_discrete(labels = x_labels)
```


```{r summarySE}
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
    library(plyr)

    # New version of length which can handle NA's: if na.rm==T, don't count them
    length2 <- function (x, na.rm=FALSE) {
        if (na.rm) sum(!is.na(x))
        else       length(x)
    }

    # This does the summary. For each group's data frame, return a vector with
    # N, mean, and sd
    datac <- ddply(data, groupvars, .drop=.drop,
      .fun = function(xx, col) {
        c(N    = length2(xx[[col]], na.rm=na.rm),
          mean = mean   (xx[[col]], na.rm=na.rm),
          sd   = sd     (xx[[col]], na.rm=na.rm)
        )
      },
      measurevar
    )

    # Rename the "mean" column    
    datac <- rename(datac, c("mean" = measurevar))

    datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean

    # Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
    ciMult <- qt(conf.interval/2 + .5, datac$N-1)
    datac$ci <- datac$se * ciMult

    return(datac)
}
```


reps1 = total_merged[tech_rep == 1]
reps2 = total_merged[tech_rep == 2]


```{r summarized_plot_bioreps}
#looking at total growth rate

into_a_plot = total_merged

x_labels_A <- addline_format(c("ColE1, pL_lac0",
                             "Sc101, pBAD"))




a <- ggplot(into_a_plot, aes(x= Vector_type,y=mu_hourly, color = Vector_type)) +
  geom_jitter(width = .2, size = 3, alpha = 0.5) +
  geom_violin(trim=FALSE, alpha = .7, draw_quantiles = c(0.5)) +
  #geom_boxplot(position = position_dodge(0.7), width = 0.1, alpha = 0.8,
  #             outlier.shape = NA) + 
  theme_linedraw() +
  theme(axis.text.y = element_text(size = 12), axis.text.x = element_text(size = 12),
        axis.title.x = element_blank(),
        legend.position = "none") +
  scale_y_continuous(breaks = c(1, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6)) +
  labs(
    x = "Retron Vector",
    y = "Growth Rate (hour^-1)",
    color = "Retron vector") +
  scale_x_discrete(labels = x_labels_A) +
  stat_compare_means(comparisons = list( c("ColE1, pL_lac0", "Sc101, pBAD")), method = "t.test", label = "p.val")


x_labels_B <- addline_format(c("GFP, Control","Retron, Cassettes",
                               "GFP, Control","Retron, Cassettes"))

b <- ggplot(into_plot, aes(x=interaction(cassette,Vector_type),y=mu_normalGFP, color = Vector_type)) +
  geom_jitter(width = .2, size = 3, alpha = 0.5) +
  geom_violin(trim=FALSE, alpha = .7, draw_quantiles = c(0.5)) +
  #geom_boxplot(position = position_dodge(0.7), width = 0.1, alpha = 0.8,
  #             outlier.shape = NA) + 
  theme_linedraw() +
  theme(axis.title.x = element_blank(), #
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        legend.position = "none") +
  scale_y_continuous(breaks = c(.7, .8, .9, 1, 1.1)) +
  labs(y = "Growth Rate (Normalized to GFP control plasmid)",
    color = "Retron vector") +
  scale_x_discrete(labels = x_labels_B) +
  stat_compare_means(comparisons = list( c(1,2), c(3,4)),
                     method = "t.test", label= "p.val")


#now, separate for all donors
addline_format <- function(x,...){
  gsub(',','\n',x)
}
x_labels <- addline_format(c("rpoB 1721,C>A",
                             "rpoB 1546,G>T",
                             "rpoB 1714,A>C",
                             "gyrA 248,C>T",
                             "rpoB 1594,delete12bp",
                             "Scrambled sequence",
                             "rpoB 443,A>T",
                             "gyrA 248,C>T (Reverse-complement)",
                             "rpoB 1547,A>G",
                             "lacZ,KO",
                             "rpoB 1538,A>C",
                             "GFP Control",
                             "rpoB 1535,C>T"))

into_plot = total_merged
c <- ggplot(into_plot, aes(x=reorder(desc,mu_normalGFP),y=mu_normalGFP, color = Vector_type)) +
  geom_boxplot(position = position_dodge(0.7), width = .5, alpha = .7 ) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1, size = 12), # 
        axis.text.y = element_text(size = 12),
        panel.grid.minor.x=element_blank(),
        panel.grid.minor.y=element_blank()) +
  scale_y_continuous(breaks = c(.7, .8, .9, 1, 1.1)) +
  labs(
    x = "Retron Donor Sequence",
    y = "Growth Rate (normalized to GFP control)",
    color = "Retron vector") +
  scale_x_discrete(labels = x_labels)

S1AB <- plot_grid(
  plotlist=list(a, b),
  align = 'v',
  axis='tblr',
  rel_widths = c(1, 1.5))

#save AB plot
ggsave(here::here('supplemental_figures/fig_supp1AB.pdf'), plot=S1AB, width=8, 
       height=4, units='in', dpi=300)
#save C plot
ggsave(here::here('supplemental_figures/fig_supp1C.pdf'), plot=c, width=8, 
       height=4, units='in', dpi=300)

S1ABC <- plot_grid(
  plotlist=list(a, b, c),
  align = 'v',
  axis='tblr',
  rel_widths = c(1, 1.5, 5))



into_b_plot = bio_reps_only
ggplot(into_b_plot, aes(x=interaction(Vector_type, donor_type),y=mu_normalGFP, color = Vector_type)) +
  geom_jitter(width = .2, size = 3, alpha = 0.5) +
  geom_violin(trim=FALSE, alpha = .7) +
  geom_boxplot(position = position_dodge(0.7), width = 0.1, alpha = 0.8,
               outlier.shape = NA) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1, size = 12), # 
        axis.text.y = element_text(size = 12)) +
  scale_y_continuous(breaks = c(.7,.8,.9, 1, 1.1)) +
  labs(
    x = "Retron Donor Sequence",
    y = "Growth Rate (normalized to GFP control)",
    title = "Effect of retron vector on growth rate",
    color = "Retron vector") 
  #scale_x_discrete(labels = x_labels)



```


```{r another plot}

#### plot some shit:
#list statistical comparisons
my_comparisons <- list( c("C_1721_A", "gfp/yfp"),
                        c("G_1546_T", "gfp/yfp"),
                        c("A_1714_C", "gfp/yfp"),
                        c("scramble/empty", "gfp/yfp"),
                        c("gyrA_248C_T", "gfp/yfp"),
                        c("1594delete12", "gfp/yfp"),
                        c("A_443_T", "gfp/yfp"),
                        c("lacZ_KO", "gfp/yfp"),
                        c("gyrA_248C_T_OTHERSTRAND", "gfp/yfp"),
                        c("rpoB1547A.G","gfp/yfp"),
                        c("C_1535_T","gfp/yfp"),
                        c("A_1538_C","gfp/yfp"))

#optional filter statements
#into_plot <- total_merged[total_merged$comments != 'very late growth?', ]
#into_plot[, mu_allele := mean(mu_normal), by=c('hard_desc')] #mean of mu, for plotting, coloring

#into_plot = total_merged


#set 1
set1 <- total_merged[total_merged$set == 1 ]
set1_summary <- ddply(set1, "desc", summarise,
                    N    = sum(!is.na(mu_normalGFP)),
                    median = median(mu_normalGFP, na.rm=TRUE),
                    mean = mean(mu_normalGFP, na.rm=TRUE),
                    sd   = sd(mu_normalGFP, na.rm=TRUE),
                    se   = sd / sqrt(N)
)
# one <- 
one <- ggplot(set1, aes(x=reorder(desc,mu_normalGFP),y=mu_normalGFP )) +
  #geom_errorbar(data = set1_summary, ymin = (mean - se), ymax = (mean + se) ) +
  geom_jitter(height = 0.0, width = 0.2, size = 2) + 
  # set color
  #color = well +
  #stat_summary(fun.data = mean_cl_normal, geom = "errorbar", fun.args = list(mult = 1)) +
  #flip interaction labels so they are vertical, aligned
  #scale_color_viridis(discrete=TRUE) +
  theme_bw() +
  theme(axis.text.x = element_text(angle =60, vjust = 1, hjust=1, size = 12),
        axis.text.y = element_text(size = 12)) +
  scale_y_continuous(breaks = c(.5, 1, 1.5, 2)) +# + 
  
  #facet_wrap(~ set) #,trans = log2_trans(),) +
  #scale_color_gradient(low="blue", high="red") +
  #scale_color_brewer(palette = "Spectral") +
  #scale_color_distiller(palette = "Spectral") +
  stat_compare_means(comparisons = my_comparisons, method = "t.test")

```

#### plot some shit:
#list statistical comparisons
my_comparisons <- list( c("C_1721_A", "gfp/yfp"),
                        c("G_1546_T", "gfp/yfp"),
                        c("A_1714_C", "gfp/yfp"),
                        c("scramble/empty", "gfp/yfp"),
                        c("gyrA_248C_T", "gfp/yfp"),
                        c("1594delete12", "gfp/yfp"),
                        c("A_443_T", "gfp/yfp"),
                        c("lacZ_KO", "gfp/yfp"),
                        c("gyrA_248C_T_OTHERSTRAND", "gfp/yfp"),
                        c("rpoB1547A.G","gfp/yfp"),
                        c("C_1535_T","gfp/yfp"),
                        c("A_1538_C","gfp/yfp"))

#optional filter statements
#into_plot <- total_merged[total_merged$comments != 'very late growth?', ]
#into_plot[, mu_allele := mean(mu_normal), by=c('hard_desc')] #mean of mu, for plotting, coloring

#into_plot = total_merged


#set 1
set1 <- total_merged[total_merged$set == 1 ]
set1_summary <- ddply(set1, "desc", summarise,
                    N    = sum(!is.na(mu_normalGFP)),
                    median = median(mu_normalGFP, na.rm=TRUE),
                    mean = mean(mu_normalGFP, na.rm=TRUE),
                    sd   = sd(mu_normalGFP, na.rm=TRUE),
                    se   = sd / sqrt(N)
)
# one <- 
one <- ggplot(set1, aes(x=reorder(desc,mu_normalGFP),y=mu_normalGFP )) +
  #geom_errorbar(data = set1_summary, ymin = (mean - se), ymax = (mean + se) ) +
  geom_jitter(height = 0.0, width = 0.2, size = 2) + 
  # set color
  #color = well +
  #stat_summary(fun.data = mean_cl_normal, geom = "errorbar", fun.args = list(mult = 1)) +
  #flip interaction labels so they are vertical, aligned
  #scale_color_viridis(discrete=TRUE) +
  theme_bw() +
  theme(axis.text.x = element_text(angle =60, vjust = 1, hjust=1, size = 12),
        axis.text.y = element_text(size = 12)) +
  scale_y_continuous(breaks = c(.5, 1, 1.5, 2)) +# + 
  
  #facet_wrap(~ set) #,trans = log2_trans(),) +
  #scale_color_gradient(low="blue", high="red") +
  #scale_color_brewer(palette = "Spectral") +
  #scale_color_distiller(palette = "Spectral") +
  stat_compare_means(comparisons = my_comparisons, method = "t.test")


#plot for set 2
set2 <- total_merged[total_merged$set == 2 ]

#set1_summary <- ddply(set1, "desc", summarise,
#                      N    = sum(!is.na(mu_normalGFP)),
#                      median = median(mu_normalGFP, na.rm=TRUE),
#                      mean = mean(mu_normalGFP, na.rm=TRUE),
#                      sd   = sd(mu_normalGFP, na.rm=TRUE),
#                      se   = sd / sqrt(N)
#)

two <- ggplot(set2, aes(x=reorder(desc,mu_normalGFP),y=mu_normalGFP )) +
    geom_boxplot() +
    geom_jitter(height = 0.0, width = 0.2, size = 2) + 
    # set color
    #color = well +
    #stat_summary(fun.data = mean_cl_normal, geom = "errorbar", fun.args = list(mult = 1)) +
    #flip interaction labels so they are vertical, aligned
    #scale_color_viridis(discrete=TRUE) +
    theme_bw() +
    theme(axis.text.x = element_text(angle =60, vjust = 1, hjust=1, size = 12),
          axis.text.y = element_text(size = 12)) +
    scale_y_continuous(breaks = c(.5, 1, 1.5, 2))  +
    stat_compare_means(comparisons = my_comparisons, method = "t.test")
  
#grid
grid.arrange(one, two, nrow = 1)



#######
#version with facet wrap:
into_plot = total_merged
ggplot(into_plot, aes(x=reorder(desc,mu_normalGFP),y=mu_normalGFP, color= interaction(tech_rep, bio_rep))) +
  geom_errorbar(height = 0.0, width = 0.2, size = 2) + 
  # set color
  #color = well +
  #stat_summary(fun.data = mean_cl_normal, geom = "errorbar", fun.args = list(mult = 1)) +
  #flip interaction labels so they are vertical, aligned
  #scale_color_viridis(discrete=TRUE) +
  theme_bw() +
  theme(axis.text.x = element_text(angle =60, vjust = 1, hjust=1, size = 12),
        axis.text.y = element_text(size = 12)) +
  scale_y_continuous(breaks = c(.5, 1, 1.5, 2)) +
  facet_wrap(~set)
  
  
  

  
#try coord flip()
ggplot(into_plot, aes(x=reorder(desc,mu_normalGFP),y=mu_normalGFP, color = as.factor(set))) +
  geom_boxplot(position = position_dodge(0.7), width = 0.5) + 
  theme_bw() +
  #theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1, size = 12), # 
  #      axis.text.y = element_text(size = 12)) +
  scale_y_continuous(breaks = c(.5, 1, 1.5, 2)) +
  labs(
    x = "Retron Donor Sequence",
    y = "Growth Rate (normalized to GFP control)",
    title = "Effect of retron vector on growth rate",
    color = "Retron vector") +
  scale_x_discrete(labels = x_labels) +
  coord_flip()

  
  
  
  
  
#optional: inspect points with ggplotly
ggplotly()

summarya27 <- ddply(into_plot, "hard_desc", summarise,
                    N    = sum(!is.na(mu_normal)),
                    median = median(mu_normal, na.rm=TRUE),
                    mean = mean(mu_normal, na.rm=TRUE),
                    sd   = sd(mu_normal, na.rm=TRUE),
                    se   = sd / sqrt(N)
)

write_path = "/Users/Max_Schubert/Dropbox/RLR_e.coli_dropbox/NewNotebook_expts/A_phenotype_in_plates/a27_more_points"
write.csv(summarya27,paste(write_path, 'summarya27.csv', sep = "/"))

annotated_dt_summ <- ddply(annotated_dt, c("genotype","locus"), summarise,
                           N    = sum(!is.na(edited.frac)),
                           median = median(edited.frac, na.rm=TRUE),
                           sd   = sd(edited.frac, na.rm=TRUE),
                           se   = sd / sqrt(N)
)



#shiiiiiiiitttttttt can I get new gp3 to better match?
#took these again with matlab, with range set as .1-.9
more_path = "/Users/Max_Schubert/Documents/MATLAB/analyze_plate_reader_growth-master"


gp3a_more = paste(more_path,"GP3_A.csv", sep = "/" )
gp3b_more = paste(more_path,"GP3_B_2.csv", sep = "/" )
gp3a_more.dt = data.table(fread(gp3a_more))
gp3b_more.dt = data.table(fread(gp3b_more))
gp3a_more.dt$plate = 3
gp3b_more.dt$plate = 3
gp3a_more.dt$rep = "a"
gp3b_more.dt$rep = "b"
combined_more = rbind(gp3a_more.dt, gp3b_more.dt)
more_merged <- combined_more[metadata, on = c("plate", "well"), nomatch = 0]

more_merged[, mean_WT := mean(mu_hourly[hard_desc == "WT"]), by=c('plate', 'rep')]
more_merged[, mu_normal := mu_hourly/mean_WT, by=c('plate', 'well')]

